---
title: "SVI and vaccinations in Wisconsin"
author: "Samantha Fabros Lee, Isha Padmanaban, Bria Thedore, and Corey Jackson"
date: "`r Sys.time()`"
output:
  html_document:
    fig_caption: yes
    highlight: pygments
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float:
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '5'
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---
# Motivation

## Research questions

# Methodology

# Analysis 

```{r, setup, include=FALSE}
library(readr) # For importing csv file
library(stargazer) # Reporting tables

# Import data set. This code chunk is where you should complete any data munging. You should import/clean data here, rename columns etc. Line 29 starts the code block and line 33 ends the code block. Notice on line 29 include=FALSE will hide code from the document that is rendered Knit the document.

 
svi_data <- read_csv("/Users/samanthalee/Documents/research/svi_clean_output.csv")
svi_data <- data.frame(svi_data)

# You might want to clean column names so they're obvious. Try to maintain consistency.
names(svi_data)[3:4] <- c("Total_Pop","Poverty_Pop")
print(svi_data)
# Extract census tract from location.
svi_data$TRACT <- gsub('^.*Tract\\s*|\\s*,[a-zA-Z]*,.*$', '', svi_data$LOCATION)
print(svi_data$TRACT)
Tot_Pop <- svi_data$Total_Pop
Poverty_Pop <- svi_data$Poverty_Pop
print(Tot_Pop)
```

You can add text here. If you want to reference data in text you can do so by placing it between ticks. Our research project has `r 1+3` people. The output file will display the number 4. 

```{r setup_analysis, include=FALSE}
# Typically we want to report high-level analysis of the data to provide the reader a window into the data. You should report how many records are in the data field. You can also report high-level statistics e.g., a table of race/ethnicity groups. 

len <- length(unique(svi_data$TRACT))
print(len)


# Create ggplot histograms for each population variable in svi_data. Add the figures in a matrix grid e.g., the code grid.arrange(p1, p2, nrow = 1) will place to figures named p1 and p2 side by side. Have 4 visualizations per row.
#install.packages("ggplot")
library(ggplot2)
#install.packages("gridExtra")
library(gridExtra)
png(file = "histogram.png")
p1 <- ggplot(svi_data, aes(Tot_Pop)) + 
  geom_histogram() +
  xlab("Total Population Overall") +
  ylab("Count") +
  ggtitle("Total Population vs. Count")



p2 <- ggplot(svi_data, aes(Poverty_Pop)) + 
  geom_histogram() +
  xlab("Below Poverty as a Socioeconomic status") +
  ylab("Count") +
  ggtitle("Poverty Population vs. Count")

Persons_aged_65_and_older <- svi_data$E_AGE65
p3 <- ggplot(svi_data, aes(Persons_aged_65_and_older)) + 
  geom_histogram() +
  xlab("Household Composition for 65 or older") +
  ylab("Count") +
  ggtitle("Persons 65 or Older vs. Count")

Persons_aged_17_and_younger <- svi_data$E_AGE17
p4 <- ggplot(svi_data, aes(Persons_aged_17_and_younger)) + 
  geom_histogram() +
  xlab("Household Composition for 17 or younger") +
  ylab("Count") +
  ggtitle("Persons 17 or Younger vs. Count")
PlotsList<- list(p1,p2,p3,p4)
grid.arrange(grobs = PlotsList, ncol = 2)
dev.off()


png(file = "histogram1.png")
Civilian_Noninstitutionalizes_with_Disability <- svi_data$E_DISABL
p5 <- ggplot(svi_data, aes(Civilian_Noninstitutionalizes_with_Disability)) + 
  geom_histogram() +
  xlab("civilians ages 5&up w/ disability") +
  ylab("Count") +
  ggtitle("Ages 5&up w/ disability vs. Count")

Single_parent_with_children_under_18 <- svi_data$E_SNGPNT
p6 <- ggplot(svi_data, aes(Single_parent_with_children_under_18)) + 
  geom_histogram() +
  xlab("Single-Parent Households w/ 18&younger") +
  ylab("Count") +
  ggtitle("Single-Parent Households vs. Count")
  

Minority_except_white_non_Hispanic <- svi_data$E_MINRTY
p7 <- ggplot(svi_data, aes(Minority_except_white_non_Hispanic)) + 
  geom_histogram() +
  xlab("Minority in Minority Status") +
  ylab("Count") +
  ggtitle("Minority vs. Count")

Speak_English_less_than_well <- svi_data$E_LIMENG
p8 <- ggplot(svi_data, aes(Speak_English_less_than_well)) + 
  geom_histogram() +
  xlab("Speaks English 'Less than Well'") +
  ylab("Count") +
  ggtitle("Minority Language vs. Count")
PlotsList<- list(p1,p2,p3,p4)
grid.arrange(grobs = PlotsList, ncol = 2)
dev.off()


png(file = "histogram2.png")
Housing_10_or_more_units <- svi_data$E_MUNIT
p9 <- ggplot(svi_data, aes(Housing_10_or_more_units)) + 
  geom_histogram() +
  xlab("Multi-Unit Structures") +
  ylab("Count") +
  ggtitle("Multi-Unit Structures vs. Count")

Mobile_homes_estimate <- svi_data$E_MOBILE
p10 <- ggplot(svi_data, aes(Mobile_homes_estimate)) + 
  geom_histogram() +
  xlab("Mobiles Homes estimate") +
  ylab("Count") +
  ggtitle("Mobiles Homes vs. Count")

More_people_than_rooms <- svi_data$E_CROWD
p11 <- ggplot(svi_data, aes(More_people_than_rooms)) + 
  geom_histogram() +
  xlab("Crowding within rooms") +
  ylab("Count") +
  ggtitle("Crowding vs. Count")

Households_with_no_vehicle <- svi_data$E_NOVEH
p12 <- ggplot(svi_data, aes(Households_with_no_vehicle)) + 
  geom_histogram() +
  xlab("Households without vehicles") +
  ylab("Count") +
  ggtitle("Transportation:no vehicle vs. Count")
PlotsList<- list(p1,p2,p3,p4)
grid.arrange(grobs = PlotsList, ncol = 2)
dev.off()



```

In total, there are `r length(unique(svi_data$FIPS))` FIPS designations in the State of Wisconsin.

```{r setup_analysis_report}
 # Display figures here. For instance you can just write p1 to display the chart or use grid.arrange(p1, p2, nrow = 1)

grid.arrange(p1,p2,p3,p4,nrow = 2)
grid.arrange(p5,p6,p7,p8,nrow = 2)
grid.arrange(p9,p10,p11,p12,nrow = 2)

```

## RQ1: What factors determine a FIPS area's socially vulnerable index score
```{r, rq1, include=FALSE}
# Here, you should conduct analyses for RQ1. The question appears to me to be a simple regression with the dependent variable RPL_Themes and independent variables all other attributes describing FIPS. 
input <- svi_data[,c("RPL_THEMES", "RPL_THEME1", "RPL_THEME2", "RPL_THEME3", "RPL_THEME4")]
model <- lm(RPL_THEMES~RPL_THEME1+RPL_THEME2+RPL_THEME3+RPL_THEME4, data = input)
model1 <- lm(RPL_THEMES~Total_Pop+Poverty_Pop, data = svi_data)
print(model1)
print(model)
RPL_THEMES_Coef <- coef(model)[1]
print(RPL_THEMES_Coef)

RPL_THEME1_Coef <- coef(model)[2]
RPL_THEME2_Coef <- coef(model)[3]
RPL_THEME3_Coef <- coef(model)[4]
RPL_THEME4_Coef <- coef(model)[5]

print(RPL_THEME1_Coef)
print(RPL_THEME2_Coef)
print(RPL_THEME3_Coef)
print(RPL_THEME4_Coef)
```

```{r, rq1_report, results='asis'}
# Report analyis for RQ1 here. Typical regression. Use the stargazer() function to report table. 
stargazer(model1,type = "html")
stargazer(model, type = "html")
```

[Add some synthesis of the regression results. What factor are significant predictors. What might this mean for vaccinations and where decision makers should place encouragement strategies??]

Total Population did not significantly predicted social vulnerability (β = -0.00005,p<0.05)

Poverty Population significantly predicted social vulnerability (β = 0.0004,p<0.05)

Socioeconomic Status significantly predicted social vulnerability (β = 0.447,p<0.05)

Household Composition and Disability significantly predicted social vulnerability (β = 0.272,p<0.05)

Minority Status and Language significantly predicted social vulnerability (β = 0.231,p<0.05)

Housing Type and Transportation significantly predicted social vulnerability (β = 0.338,p<0.05)
